### 2009-08-13 doc/tech_doc/sec/arc-security-documentation.tex
more changes for sec documentation

It's configuration is presented in Section 7.5.

@@

\subsection{UsernameToken SecHandler} % (fold)
\label{subsec:username_token}
The UsernameToken SecHandler is meant for processing - generating and extracting - WS-Security [4] UsernameToken in the SOAP header. Hence it must be attached to the MCC which processes SOAP payloads - like SOAP MCC or Service accepting SOAP messages. For description of configuration see section 8.13.

@@

For the outgoing message this SecHandler creates adds proper token into SOAP header according to configuration.

\subsection{X509Token SecHandler} % (fold)
\label{subsec:x509_token}
The X509Token SecHandler is meant for processing - generating and extracting - WS-Security [4] X509Token from SOAP header. Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages. For description of configuration see section 8.14.

@@

\subsection{SAMLToken SecHandler} % (fold)
\label{subsec:saml_token}

@@

@@

@@

\subsection{Interface of policy decision service} % (fold)
\label{subsec:interface_pds}
WSDL with comments available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.wsdl .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/mcc/tls/tls.xsd .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/pdp/pdp.xsd .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcauthzsh/SimpleListAuthZ.xsd.

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/simplelistpdp/SimpleListPDP.xsd .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/arcpdp/ArcPDP.xsd .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/pdpserviceinvoker/ArcPDPServiceInvoker.xsd .

@@

\begin{verbatim}
<Service name="echo" id="echo">
  <SecHandler name="arc.authz" id="authz" event="incoming">
    <!--Remote pdp service invoking-->
      <PDP name="pdpservice.invoker">
        <ServiceEndpoint>https://127.0.0.1:60001/pdp.service</ServiceEndpoint>
        <KeyPath>./key.pem</KeyPath>
        <CertificatePath>./cert.pem</CertificatePath>
        <CACertificatePath>./ca.pem</CACertificatePath>
      </PDP>
  </SecHandler>
  <next id="echo"/>
  <echo:prefix>[ </echo:prefix>
  <echo:suffix> ]</echo:suffix>
</Service>
\end{verbatim}

@@

The PDP Service Invoker is a client of PDP Service. The configuration options include endpoint of service and credentials to be used for establishing secure connection.

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/delegationsh/DelegationPDP.xsd .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/usernametokensh/UsernameTokenSH.xsd .

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/pdc/x509tokensh/X509TokenSH.xsd.

@@

@@

\section{SAML Attribute Authority Service} % (fold)
\label{sec:saml_aa_service}

>>>>

It's configuration is presented in Section \ref{subsec:pdpservice_conf}.

@@

\subsection{UsernameToken SecHandler} % (fold)
\label{subsec:username_token}
The UsernameToken SecHandler is meant for processing - generating and extracting - WS-Security [4] UsernameToken in the SOAP header. Hence it must be attached to the MCC which processes SOAP payloads - like SOAP MCC or Service accepting SOAP messages. For description of configuration see section \ref{subsec:ut_sechandler_conf}.

@@

For the outgoing message this SecHandler creates and adds proper token into SOAP header according to configuration.

\subsection{X509Token SecHandler} % (fold)
\label{subsec:x509_token}
The X509Token SecHandler is meant for processing - generating and extracting - WS-Security [4] X509Token from SOAP header. Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages. For description of configuration see section \ref{subsec:xt_sechandler_conf}.

@@

\subsection{SAMLToken SecHandler} % (fold)
\label{subsec:saml_token}
The SAMLToken SecHandler is meant for processing - generating and extracting - WS-Security [4] SAMLToken from SOAP header. Hence it must be attached to the MCC which processes SOAP payloads – like SOAP MCC or Service accepting SOAP messages. For description of configuration see section \ref{subsec:st_sechandler_conf}.

For the outgoing message this SecHandler creates SAMLToken in SOAP header. SOAP message body is encrypted and signed.

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Policy.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Request.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Response.xsd

\subsection{Interface of policy decision service (Charon service)} % (fold)
\label{subsec:interface_pds}
WSDL with comments available at

The following is the configuration about the Charon service, which is configured to use xacml policy engine. The policy engine is configurable by changing the ``name'' attribute of three elements: \textit{$<$Evaluator$>$}, \textit{$<$Policy$>$}, \textit{$<$Request$>$} into ``arc.evaluator'', ``arc.policy'', ``arc.request''. Also the policy should also be changed to the one with ARC specific format.

\begin{verbatim}
    <Service name="charon" id="charon_service">
        <!--The element <Evaluator/>, <Policy/> and <Request/> configuration
        are supposed to be used to load object; element <PolicyStore/> is
        supposed to be used to get the location of policy-->
        <charon:PDPConfig>
            <charon:PolicyStore>
                <charon:Location Type="file">charon_policy_xacml.xml.example</charon:Location>
                <!-- other policy location-->
            </charon:PolicyStore>
            <charon:Evaluator name="xacml.evaluator" />
            <charon:Policy name="xacml.policy" />
            <charon:Request name="xacml.request" />
        </charon:PDPConfig>
    </Service>
\end{verbatim}

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/mcc/tls/tls.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/services/charon/charon.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcauthzsh/SimpleListAuthZ.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/simplelistpdp/SimpleListPDP.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/ArcPDP.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/pdpserviceinvoker/PDPServiceInvoker.xsd

@@

\begin{verbatim}
<Service name="echo" id="echo">
  <SecHandler name="arc.authz" id="authz" event="incoming">
    <!--Remote pdp service invoking-->
      <PDP name="pdpservice.invoker">
        <ServiceEndpoint>https://127.0.0.1:60001/pdp.service</ServiceEndpoint>
        <KeyPath>./key.pem</KeyPath>
        <CertificatePath>./cert.pem</CertificatePath>
        <CACertificatePath>./ca.pem</CACertificatePath>
        <RequestFormat>XACML</RequestFormat>
        <TransferProtocol>SAML</TransferProtocol>
      </PDP>
  </SecHandler>
  <next id="echo"/>
  <echo:prefix>[ </echo:prefix>
  <echo:suffix> ]</echo:suffix>
</Service>
\end{verbatim}

@@

The PDP Service Invoker is a client of PDP Service. The configuration options include endpoint of service and credentials to be used for establishing secure connection. In addition, the $<$RequestFormat$>$ element is for specifying the format of the request, and the $<$TransferProtocol$>$ element is for specifying the protocol of tranfering the request.

Table \ref{table:support_pdp_request_protocol} shows the support of request and protocol in remote policy decision making. Note that if ``SAML 2.0 profile of XACML v2.0'' is configured in ``pdpservice invoker'', besides the authorizaton service (called charon service) implemented in ARC, it can interact with external authorization services, such as gLite authorization service (Yet the interoperation test has not been done).

\begin{table}[ht]
\caption{Support of request and protocol in remote policy decision making}
\centering
\begin{tabular}{| l | p{7cm} | p{5cm} |}
\hline
\textbf{  } & \textbf{ARC Request/Response} & \textbf{XACML Request/Response} \\ \hline
\textbf{ARC protocol} & support & support \\ \hline
\textbf{SAML protocol} & not support & support (SAML 2.0 profile of XACML v2.0) \\ \hline
\end{tabular}
\label{table:support_pdp_request_protocol}
\end{table}

\textbf{ARC protocol}:

\textbf{SAML protocol}:

http://www.oasis-open.org/committees/download.php/11475/access\_control-xacml-2.0-saml-protocol-schema-os.xsd

\textbf{ARC request/response}:

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Request.xsd

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/arcpdp/Response.xsd

\textbf{Note} that the format of ``Response'' is supposed to correspond with the format of ``Request''.

\textbf{XACML request/response}:

http://www.oasis-open.org/committees/download.php/11474/access\_control-xacml-2.0-saml-assertion-schema-os.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/delegationpdp/DelegationPDP.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/usernametokensh/UsernameTokenSH.xsd

@@

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/x509tokensh/X509TokenSH.xsd

@@

\subsection{SAMLToken SecHandler Configuration} % (fold)
\label{subsec:st_sechandler_conf}
Configuration XML schema with comments is available at

http://svn.nordugrid.org/trac/nordugrid/browser/arc1/trunk/src/hed/shc/samltokensh/SAMLTokenSH.xsd

@@

\section{SAML Attribute Authority Service} % (fold)
\label{sec:saml_aa_service}
SAML Attribute Authority Service (SAML AA Service) is implemented for accepting SAML \textit{$<$samlp:AttributeQuery$>$} from the requester, and issuing SAML \textit{$<$samlp:Response$>$} which is supposed to include \textit{$<$saml:Assertion$>$} that contains the attribute information about the requester.

SAML AA service is implemented with the same functionality as VOMS SAML service. More specifically, the interface of SAML AA service is the same as VOMS SAML service.

The client functionality of \textbf{SAML AA service} exists in two places:

\begin{itemize}
    \item The specific client utility (saml\_assertion\_init), where the SAML AA service is contacted to get back the SAML assertion.

    \item The security handler about SAML token, where the SAML AA service is contacted and then the responded SAML assertion is inserted into SOAP message according to WS-Security SAML Token profile.

The same as VOMS SAML Service, the VOMS Admin service is used in ARC for managing the attributes. Therefore the database schema from VOMS is also introduced into ARC.

However the implementation of SAML AA Service in ARC is designed not to be coupled with any specific backend database. Therefore even if the database schema is changed, the implementation is not necessarily going to be changed. The way about decoupling the service implementation with database schema is through the service configuration.

\begin{verbatim}
        <Service name="aa.service" id="aa_service">
            <aa:KeyPath>../../tests/echo/testkey-nopass.pem</aa:KeyPath>
            <aa:CertificatePath>../../tests/echo/testcert.pem</aa:CertificatePath>
            <aa:CACertificatesDir>../../tests/echo/certificates</aa:CACertificatesDir>
            <aa:CACertificatePath>../../tests/echo/testcacert.pem</aa:CACertificatePath>

            <aa:Database type="mysql" ip="localhost" port="3305" dbname="voms_knowarc" user="root" 
              password="123456">
              <!--The sql following sentences and their combination are introduced
              according to the code of voms' mysql plugin.
              The intention is that if some other database schema is used, you can
              simply change the sql sentences here, without changing the code. But still,
              the input argument to these queries should be with the same meaning,
              and the same sequence as the following; also the output of these queries
              should also be with the same meaning and the same sequence as the following-->
              <aa:SQLSet name="UID">
                <aa:SQL name="GetUID">SELECT userid FROM usr WHERE usr.dn = ?</aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="Role">
                <aa:SQL name="GetRole">SELECT groups.dn, role FROM groups, 
                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
                  AND m.userid = ? AND roles.role = ?</aa:SQL>
                <aa:SQL name="GetGroup">SELECT groups.dn, NULL FROM groups, m 
                  WHERE groups.gid = m.gid AND m.userid = ?</aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="Group">
                <aa:SQL name="GetGroup">SELECT groups.dn, NULL FROM groups, m 
                  WHERE groups.gid = m.gid AND m.userid = ?</aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="GroupAndRole">
                <aa:SQL name="GetGroup">SELECT groups.dn, NULL FROM groups, m 
                  WHERE groups.gid = m.gid AND m.userid = ?</aa:SQL>
                <aa:SQL name="GetGroupAndRole">SELECT groups.dn, role FROM groups, 
                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
                  AND m.userid = ? AND roles.role = ? AND groups.dn = ?</aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="All">
                <aa:SQL name="GetAll">SELECT groups.dn, role FROM groups, 
                  m LEFT JOIN roles ON roles.rid = m.rid WHERE groups.gid = m.gid 
                  AND m.userid = ?</aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="GroupAndRoleAttribute">
                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
                </aa:SQL>
                <aa:SQL name="GetGroupAttribute">SELECT attributes.a_name, 
                  group_attrs.a_value, groups.dn, NULL FROM attributes, group_attrs, 
                  groups, m WHERE attributes.a_id = group_attrs.a_id AND groups.gid = m.gid 
                  AND m.userid = ? AND m.rid is NULL AND group_attrs.g_id = m.gid
                </aa:SQL>
                <aa:SQL name="GetGroupAndRoleAttribute">SELECT attributes.a_name, 
                  role_attrs.a_value, groups.dn, roles.role FROM attributes, role_attrs, 
                  groups, roles, m WHERE attributes.a_id = role_attrs.a_id 
                  AND groups.gid = m.gid AND m.userid = ? AND m.rid = roles.rid 
                  AND roles.role = ? AND groups.dn = ? AND role_attrs.g_id = m.gid 
                  AND role_attrs.r_id = m.rid
                </aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="GroupAttribute">
                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
                </aa:SQL>
                <aa:SQL name="GetGroupAttribute">SELECT attributes.a_name, 
                  group_attrs.a_value, groups.dn, NULL FROM attributes, group_attrs, 
                  groups, m WHERE attributes.a_id = group_attrs.a_id AND groups.gid = m.gid 
                  AND m.userid = ? AND m.rid is NULL AND group_attrs.g_id = m.gid
                </aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="RoleAttribute">
                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
                </aa:SQL>
                <aa:SQL name="GetRoleAttribute">SELECT attributes.a_name, 
                  role_attrs.a_value, groups.dn, roles.role FROM m 
                  INNER JOIN groups ON m.gid = groups.gid LEFT JOIN roles 
                  ON roles.rid = m.rid INNER JOIN role_attrs on groups.gid = role_attrs.g_id 
                  INNER JOIN attributes on attributes.a_id = role_attrs.a_id 
                  WHERE role_attrs.r_id = roles.rid AND m.userid = ? AND roles.role = ?
                </aa:SQL>
              </aa:SQLSet>

              <aa:SQLSet name="AllAttribute">
                <aa:SQL name="GetUserAttribute">SELECT attributes.a_name, 
                  usr_attrs.a_value, NULL, NULL FROM attributes, usr_attrs 
                  WHERE attributes.a_id = usr_attrs.a_id AND usr_attrs.u_id = ?
                </aa:SQL>
                <aa:SQL name="GetGroupAttribute">SELECT attributes.a_name, 
                  group_attrs.a_value, groups.dn, NULL FROM attributes, group_attrs, 
                  groups, m WHERE attributes.a_id = group_attrs.a_id AND groups.gid = m.gid 
                  AND m.userid = ? AND m.rid is NULL AND group_attrs.g_id = m.gid
                </aa:SQL>
                <aa:SQL name="GetGroupAndRoleAttributeAll">SELECT attributes.a_name, 
                  role_attrs.a_value, groups.dn, roles.role FROM attributes, role_attrs, 
                  groups, roles, m WHERE attributes.a_id = role_attrs.a_id 
                  AND groups.gid = m.gid AND m.userid = ? AND m.rid = roles.rid 
                  AND role_attrs.g_id = m.gid AND role_attrs.r_id = m.rid
                </aa:SQL>
              </aa:SQLSet>

              <!--Other SQLSet-->

            </aa:Database>
        </Service>
\end{verbatim}
